
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("tidyverse") #This project uses tidyverse for pipeline.
```

## Data Cleaning and Visualization for CSV of Study Performance

This project uses a CSV file from Kaggle. I have modified the CSV to have some blank values to showcase data cleaning. The inital CSV file is from: https://www.kaggle.com/datasets/spscientist/students-performance-in-exams

### Data Cleaning

In this section, it will: 
  1. Clean up the data values of NA. 
  2. Removes the word of 'group' in the race_ethnicity section due to its redundancy. 

```{r Data_Cleaning}

study_performance <- read.csv("study_performance.csv", na.strings = "") #reading the csv of the study performance. I got this data set from Kaggle.

summary(study_performance) #to see the columns of the csv

head(study_performance)

######### Cleaning up data values ##########

study_performance[is.na(study_performance)] <- 0 #making the NA values to be zero
head(study_performance) #printing a few rows to check if it worked

#in the race_ethnicity, removing 'group' because it is too redunant
#reference is from: https://www.digitalocean.com/community/tutorials/sub-and-gsub-function-r

study_performance <- study_performance %>% mutate(race_ethnicity = gsub("group ", "", race_ethnicity)) #changing the word 'group' to be blank. Also using mutate for dplyr function.

head(study_performance) #printing out a few of the rows in race_ethnicity to check.


```

### Putting Data in Columns
This section will create a new column for the average score. It will also delete columns of math_score, reading_score, and writing_score.

```{r Data_Columns}

study_performance <- study_performance %>% mutate(average_score = (math_score + reading_score + writing_score) /3) #adding another column called average_score usind dyplr's mutate

study_performance <- study_performance %>% select( -c(math_score, reading_score, writing_score)) #deleting the columns of math_score, reading_score, and writing_score

head(study_performance) #printing out a few of the rows to see if it created a new column and deleted the old columns.

```


### Data Visualization
This section will create a data visualization for the data of:
  1. different race ethnicity vs the average score in a boxplot
  2. gender and race ethnicity against the average score in a plot.
  
The 2 different data visualization will use GGPLOT2.

```{r Data_Visualization}

###PLOT 1 ###
#creating a boxplot for the different race ethnicity vs the average score
ggplot(study_performance, aes(x = race_ethnicity, y = average_score, color = race_ethnicity )) +
  geom_boxplot() +
  facet_wrap(~race_ethnicity) +
  labs(x = "Race Ethnicity", y = "Average Score", title = "The Performance of Race Ethnicity Over Average Score") + theme_classic() + scale_color_brewer(type = "qual", palette = 2)


###PLOT 2 ###
#creating a plot of the gender and race ethnicity against the average score

ggplot(study_performance, aes(fill = race_ethnicity, x = gender, y = average_score)) + 
  geom_bar(position = 'dodge', stat = 'identity') +
  labs(x = "Gender", y = "Average Score", title = "The Gender and Race Ethnicity Against the Average Score") + scale_fill_manual(values = c("#7c6d9e", "#291e42", "#26408B", "#A6CFD5", "#C2E7D9"))

```
